### 2016-12-05 Limber_Comment.tex
few changes and comments

\documentclass[preprint]{aastex}
\usepackage{graphicx}   % Including figure files
\usepackage{amsmath}    % Advanced maths comm
\usepackage{color,verbatim,url}
\usepackage{tabularx}
\usepackage[table,usenames,dvipsnames]{xcolor}\usepackage{pdflscape}
\usepackage{lastpage}

\definecolor{pink}{rgb}{0.858, 0.188, 0.478}
\definecolor{purple}{RGB}{76, 0,153}

\newcommand{\red}[1]{{\color{red}{#1}}}
\newcommand{\green}[1]{{\color{green}{#1}}}
\newcommand{\blue}[1]{{\color{blue}{#1}}}
\newcommand{\x}{\vec{x}}
\renewcommand{\d}[0]{{\rm d}}
\newcommand{\shahab}[1]{\textcolor{pink}{\bf #1}}

\newcommand{\be}{\begin{equation}}  \newcommand{\ee}{\end{equation}}
\newcommand{\mb}[1]{\mbox{ #1 }}  
\newcommand{\ba}{\begin{eqnarray}}\newcommand{\ea}{\end{eqnarray}}
\newcommand{\bm}[1]{\mbox{\boldmath{$#1$}}}   %this is bold italic for MNRAS
\newcommand{\mat}[1]{\mathbfss{#1}}

\begin{document}

\title{\huge{Comment on `The Limits of Cosmic Shear'}}

\author{Members of the KiDS, CFHTLenS and RCSLenS collaborations\altaffilmark{1} ....}
%\email{tbd@tbd}

%\altaffiltext{1}{}

\begin{abstract}
Kitching et al. (2016) examine the impact of the flat-sky and Limber approximations that can be used when constraining cosmological parameters from the measurement of weak gravitational lensing by large-scale structures, also known as cosmic shear.  We support the core message of this paper that the use of such approximations should be considered in the future analyses of high precision surveys such as Euclid, LSST and WFIRST.  \shahab{(not sure it is clear that they are important even for future surveys, when considered together vs not considered at all)} In this comment, we argue that these approximations have negligible impact for current surveys and are not able to explain away the tension between the constraints from some current weak lensing surveys and Planck.  Reducing uncertainty and characterising bias in photometric redshift estimation is a critical task for the success of current and future surveys.  We highlight that sophisticated treatments of these uncertainties have already been explored in the literature and have been shown not to resolve the tension with Planck, in contrast to the message that Kitching et al. (2016) convey \shahab{(do we perhaps want to remove the last clause "in contrast to the message..." to make the draft somewhat less Kitching-centric given the weakness of that analysis?)}.

\end{abstract}
\section{Introduction}

The measurement of weak gravitational lensing by large scale structures provides a powerful cosmological probe of dark matter, dark energy, and modifications to gravity.  As such it is the primary science goal of several current (KiDS, HSC, DES) and future (Euclid, LSST, WFIRST) large-scale surveys. Interest in the results from these surveys is high as statistically significant deviations have been found between the cosmological parameter constraints from the CMB Planck experiment \citep{planck/cosmo:2015} in comparison to weak lensing constraints from both the Kilo-Degree Survey \citep[KiDS;][]{hildebrandt/etal:2016} and the Canada-France-Hawaii Telescope Lensing Survey \citep[CFHTLenS;][] {joudaki/etal:2016}.  If the source of this tension is not a result of so-far unconsidered sources of systematic errors in one or all experiments, extensions to the standard flat $\Lambda$CDM cosmological models need to be considered.  \citet{joudaki/etal:2017} have shown, for example, that the tension can be resolved with an evolving dark energy model.

In this comment paper we discuss the flat sky and Limber approximations that can be used when deriving theoretical predictions for the principally used cosmic shear statistic, $\xi_\pm$, the two-point shear correlation function. The Limber approximation for cosmic shear has been previously explored by \citet{giannantonio/etal:2012} finding that it provided sufficient accuracy for future cosmic shear studies. Recently, however, \citet{kitching/etal:2016} reported that when considering the flat sky and Limber approximations in concert, tensions between weak lensing and CMB experiments can be reduced. This was demonstrated with a re-analysis of the non-tomographic weak lensing data from the CFHTLenS survey \citep{kilbinger/etal:2013}.

This comment is structured as follows.  In section~\ref{sec:theory} we clarify which brand of approximations, highlighted by \citet{kitching/etal:2016}, have been used in recent weak lensing analyses \citep{joudaki/etal:2016, hildebrandt/etal:2016, joudaki/etal:2017}.  We show that we are unable to replicate key results from \citet{kitching/etal:2016} and therefore do not expect these approximations to impact on the cosmological analysis of current surveys. In section~\ref{sec:cfhtlens} we compare the `one-parameter' cosmological re-analysis of \citet{kitching/etal:2016} with the full six-parameter analysis by \citet{kilbinger/etal:2013}, arguing that this particular non-tomographic data set is not in significant tension with Planck (\shahab{not sure I agree with this clause, please see my email}).  We also briefly review the literature that has advanced our understanding since the original CFHTLenS analyses were published.

In section~\ref{sec:photoz} we review our current knowledge of photometric redshift uncertainties and the sophisticated mitigation strategies that have already been applied to data.  We suggest that a possible misinterpretation in the direction of the redshift biases reported in \citet{choi/etal:2016} has lead to the incorrect conclusion of \citet{kitching/etal:2016} that these redshift biases can resolve the tension with Planck.

\section{Theory}
\label{sec:theory}
\input{theory.tex}

\section{Advances in understanding since CFHTLenS}
\label{sec:cfhtlens}
\input{cfhtlens.tex}

\section{Photometric redshift bias and uncertainty}
\label{sec:photoz}
\input{photoz.tex}

\section{Conclusion}
\label{sec:conclusion}
\input{conclusion.tex}

\bibliographystyle{apj}
\bibliography{Limber_Comment}
\end{document}