### 2017-02-10 conclusions.tex
Figures updates labels

In this paper we evaluate precision theoretical calculations for cosmic shear
observables, bringing together different sources from the literature to provide
a pedagogical review of the impact of adopting flat-sky and Limber
approximations.  We demonstrate that for current surveys, such as CFHTLenS,
these approximations have a negligible impact on cosmological parameter
constraints.

For future surveys, the decrease in statistical errors places higher
requirements on the accuracy of the theoretical modelling.    There is also,
however, the need to be able to rapidly sample the multi-dimensional
cosmological parameter likelihoods.  This requirement for computational speed
is incompatible with a theoretical analysis that calculates a full spherical
solution for the shear power spectrum, without adopting any approximation.  We
therefore present alternative solutions, revisiting the work of
\citet{2012PhRvD..86b3001B} who showed that adopting the second-order extended
Limber approximation of \citet{2008PhRvD..78l3506L} provides a representation
of the full spherical solution for the shear power spectrum that is accurate at
the sub-percent level for $\ell > 3$.    We have verified this result and
provide to the community our fast numerical implementation of all the
approximations studied in this analysis, and the slow calculation of the full
projection within the publicly available package \textsc{nicaea} at
\texttt{http://www.cosmostat.org/software/nicaea}.

Finally we propose that future surveys seek to optimise the statistical
analyses of their cosmic shear data.  For example moving from the standard
two-point shear correlation function statistic to the more stringent `COSEBI'
statistic \citep{COSEBIs} renders the cosmic shear measurement insensitive to
the low-$\ell$ scales where the Limber and flat-sky approximations have an
impact on the precision of the theoretical modelling.

\mk{TODO: comment on non-flat models.}