### 2017-04-20 answer_to_referee.txt
First draft of referee reply and modifications.

We would like to thank the referee for their positive report and useful comments and suggestions.
Please fine below our replies to the referee's points. All modifications to the draft are highlighted
in red.

- Shear correlation function, eq. (47-48).
  We had indeed used a transformation (with the Legendre polynomial) that is incorrect for spin-2 correlations.
  We have replaced this with the correct transformations using the reduced Wigner D-matrices. We state their
  form in a new Appendix C. Figures 2 and 3, involving the shear correlation functions, have been updated
  accordingly.

- p. 6, chi-dependence of \tilde P_m.
  We added to the equation an approximate expression of P_m(k), without the chi-dependence. We hope that the
  subsequent text as it stands should make the conditions under which this approximation holds sufficiently clear.

- p. 7, accuracy of the ExtL2FlHyb.
  * Not sure about the reason. *

- Making Fig. 1 and following clearer.
  * Not sure that adding labels or anything else would make those clearer. *

- Table 1, extra column.
  We added columns indicating the ell_x such that for ell>ell_x the power spectrum projection has better
  accuracy than x, with x=10% and 1%.

- Fig. 3, comparison for xi_-.
  This plot has been updated, both xi_+ and xi_- are now compared to the full spherical case.

- Table 2, adding plot.
  * Really necessary? *

- Conclusions
  * I think present tense is fine *

- Fig. 4.
  * On my screen I don't have a problem with this figure. *

Other modifications.

- End of introduction: Lemos et al. reference added.

- Eq. 47: The definitions as <gamma gamma^*> and <gamma gamma> for xi_+ and xi_- have been added here,
  moved from eq. 48.

- Typos:
  All typos have been fixed.