### 2017-07-18 answer_to_referee.txt

We would like to thank the referee for their positive report and useful comments and suggestions.
Please find below our replies to the referee's points. All modifications to the draft are highlighted
in red.

- Major point, shear correlation function, eq. (47-48).
  We had indeed used a transformation (with the Legendre polynomial) that is incorrect for spin-2 correlations.
  We have replaced this with the correct transformations using the reduced Wigner D-matrices. We added a short
  Appendix C for details on their computation. Figures 2 and 3, involving the shear correlation functions, have been updated
  accordingly, as well as some of the explanatory text.

Minor points:

- abstract: Removed comment on Kitching et al. (2016, version 1), since their version 2 did not retain
  the incorrect calculations and claims from v1.

- p. 6, chi-dependence of \tilde P_m.
  We added to the equation an approximate expression of P_m(k), without the chi-dependence. We hope that the
  subsequent text as it stands should make the conditions under which this approximation holds sufficiently clear.

- p. 7, accuracy of the ExtL2FlHyb.
  In all cases, the second-order expansion adds power to the first-order term, as can be seen in Fig. 1. ExtL1FlHyb over-estimates the full spherical projection,
  and thus ExtL2FlHyb increases this over-estimation even more. This expansion would converge to a full flat case (eq. 29), which we do not plot, and there is no
  reason to believe that the difference to FullSph decreases with the expansion order.
  For ExtL1FlHyb versus L1Fl, we see the better convergence of the former
  for ell>8, as expected, and did not try to explain why L1F1 is slightly better at very low ell.

- Making Fig. 1 and following clearer.
  We could not think of straight-forward ways to improve the readibility of the plots. We hope that the combination
  of colors and line styles is sufficient to tell apart the different cases.

- Table 1, extra column.
  We added two columns indicating ell_x such that for ell>ell_x the power spectrum projection has better
  accuracy than x, with x=10% and 1%.

- Fig. 3, comparison for xi_-.
  This plot has been updated, both xi_+ and xi_- are now compared to the full spherical case.

- Table 2, adding plot.
  We believe that a plot version of Table 2 does not enhance the value of the paper, since the values of sigma_8 (Omega_m/0.27)^0.6
  are so close together, well within the 68% credible interval.

- Conclusions
  We believe that the present tense mode is appropriate for the conclusion of the paper.

- Fig. 4.
  This plot seems fine on our screens and printouts, we do not detect a problem of poor resolution.

Other small modifications.

- End of introduction: Lemos et al. reference added.

- Eq. 47: The definitions as <gamma gamma^*> and <gamma gamma> for xi_+ and xi_- have been added here,
  moved from eq. 48.

- Table 2: In the caption, we replaced the expression "confidence interval" with "credible interval". For the ExtL2Sph case,
  we added "Hankel" to clarify that we use this approximation for the correlation function.

- Typos:
  All typos have been fixed.